---
title: "Methods for SEM methods section"
author: 
output:
   word_document:
      reference_docx: working_docx_template.docx
      toc: no
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
      - '--filter=pandoc-crossref.exe'
   pdf_document:
     keep_tex: true
   html_document: default
   pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
      - '--filter=pandoc-crossref.exe'
header-includes:
- \usepackage{lineno}
- \usepackage{amsmath}
- \numberwithin{equation}
- \usepackage{indentfirst}
- \linenumbers
indent: true
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

```{r import general objects, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# read in the path analysis model syntax
source("habitat-model-specs.R")
library(knitr)
library(tidyverse)
library(lavaan)
## load the data ----
dat = read.csv("./data/divers_prod_hab.csv") %>% 
  dplyr::rename(prod = 'production_sum_mean', habH = 'shannon_habitat',
                bioH_richness = 'RICHNESS', bioH_shannon = 'SHANNON_B',
                bioH_simpson = 'SIMPSON_B', bioH_evenness = 'EVENNESS_B', is = 'median_nolog', sediment_size = 'seg_weighted_size') %>% 
  # scale the variables because production is on a vastly different scale than rest
  dplyr::mutate(across(c(prod, habH,bioH_richness, bioH_shannon, bioH_simpson, bioH_evenness, sediment_size, is), ~as.numeric(scale(.x)), .names = "{.col}_scaled")) %>% 
  dplyr::select(prod, prod_scaled, habH, habH_scaled, bioH_richness, bioH_richness_scaled, bioH_shannon, bioH_shannon_scaled, bioH_simpson, bioH_simpson_scaled, bioH_evenness, bioH_evenness_scaled,sediment_size, sediment_size_scaled, is, is_scaled)

# SEMs ----
# fit with full path analysis/sems
model_latentBioH_scaled <-'
# set out the model structure
# the full model of interaction strengths
# IS is influenced by production and diversity

is_scaled ~ prod_scaled + bioH_scaled

# here prod itself has dependence on habitat Diversity
# and biodiversity and sediment size

prod_scaled ~ habH_scaled + bioH_scaled+ sediment_size_scaled

# OTOH, biodiversity is influenced by habitat Diversity and sediment size

bioH_scaled ~ habH_scaled + sediment_size_scaled

sediment_size_scaled ~ habH_scaled

## identify the latent variables

bioH_scaled =~ bioH_shannon_scaled + bioH_richness_scaled + bioH_evenness_scaled
'

### models with latent variable for biodiversity (richness, shannon, evenness)
# sem model 
fit_sem_bioH_scaled <- sem(model_latentBioH_scaled, data = dat, auto.var = TRUE)

# extract the relevant summary stats from the fitted models
# summary(fit_sem_bioH_scaled,fit.measures = TRUE, standardized = TRUE, rsquare = TRUE)


# estimate the standardized coefficients
# standardizedsolution(fit_sem_bioH_scaled)
```
<!--- Title page details --->

We proposed habitat diversity would modulate food web module interaction strengths through its effects on prey resource diversity and production. To quantify this effect, as well as the relative importance of prey diversity and production on interaction strengths, we needed to account for the variety of direct and indirect pathways that habitat diversity influences prey resource diversity and secondary production. Therefore, we used structural equation models [SEM, @grace; @shipley] to examine the multiple causal pathways (Fig. 1) while accounting for the correlation between the multiple variables. We hypothesized habitat diversity, measured by Shannon-Weiner entropy, would influence both the Models were fit using the `lavaan` package [ver. 0.6-15, @rosseel2012]. 


```{r standard table}
knitr::kable(standardizedsolution(fit_sem_bioH_scaled), digits = 3)

```

# References

<div id="refs"></div>
